\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2020-07-24 sections/discussion.tex
Updates from Overleaf

% General statements
Co-occurrence associations in microbial communities help identify important interactions that drive microbial community structure and organization.
Our analysis shows that networks generated using different combinations of tools and approaches can look significantly different from each other, highlighting the importance of a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices.
Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
The \ac{micone} pipeline is built in a modular fashion.
Its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from those that were not examined in the present work and those that will be developed in the future.

>>>>

% General statements
Co-occurrence associations in microbial communities help identify important interactions that drive microbial community structure and organization.
Our analysis shows that networks generated using different combinations of tools and approaches can look significantly different from each other, highlighting the importance of a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices (refer to the Methods section).
Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
The \ac{micone} pipeline is built in a modular fashion.
Its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from those that were not examined in the present work and those that will be developed in the future.

### 2020-07-24 sections/results.tex
Updates from Overleaf

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.69\linewidth]{figure1.pdf}
    \caption{
      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
      The processes can be grouped into four major steps: \textbf{(DC)} denoising and clustering, \textbf{(TA)} taxonomy assignment, \textbf{(OP)} OTU/ESV processing, and \textbf{(NI)} network inference.
      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

@@

  Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
  In order to compare the assigned taxonomies from different databases, we use the same reference sequences and assign taxonomies to them using different taxonomy reference databases.
  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq.
  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
  The \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on December 2017 at the time of writing and \ac{ncbi} is updated as new sequences are curated.
  Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies~\cite{Balvociute2017}.

@@

  Given that no database performs better than others against mock datasets, and that databases are almost equally distant from each other in terms of final output, the choice of which database to use should be driven by other reason.
  One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see also Discussion). Another reason could be the frequency of updates and the potential for future growth, which prompted us to set \ac{ncbi} as the \ac{micone} standard for taxonomy assignment.
 In addition to being regularly maintained and updated the \ac{ncbi} database has the advantage that its accuracy of assignments is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.

@@

  % TODO: Talk about the difference between correlations and associations
  The six different network inference methods used in this study are \ac{magma}~\cite{Cougoul2019}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
  These network inference methods into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
  Pairwise correlation methods involve calculating the correlation coefficient between every pair of \ac{OTU}/\ac{esv}s leading to the detection of spurious indirect connections.
  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected \ac{otu}s.

  For the analysis presented in this section, we used the taxonomy composition table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
  Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
  We can clearly see that the different networks differ vastly in their edge-density and connectivity and even some of the edges in common to these networks have their signs inverted. Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on [REF]).
  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison).
  The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
  The edge intersections in contrast show that only $19$ edges (out of $98$ edges in the smallest network - \ac{magma}) are in common between all the methods and each network has a large number of unique edges.
  These results indicate that there is a substantial rewiring of connections in the inferred networks.

@@

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{figure5.pdf}
  \end{figure}
  \begin{figure}[t!]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links are negative associations.
      A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
      \textbf{(B)} The node overlap upset plot indicates that all the networks have a large number of common nodes involved in connections.
      Whereas, \textbf{(C)} The edge overlap upset plot shows that a very small fraction of these connections are actually shared.
    }
    \label{fig:figure5}
  \end{figure}

>>>>

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.69\linewidth]{figure1.pdf}
    \caption{
      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
      The steps can be grouped into four major groups: \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

@@

  Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
  In order to compare the assigned taxonomies from different databases, we use the same reference sequences and assign taxonomies to them using different taxonomy reference databases.
  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq.
  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
  The \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on December 2017 at the time of writing and \ac{ncbi} is updated as new sequences are curated.
  Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.

@@

  Given that no database performs better than others against mock datasets, and that databases are almost equally distant from each other in terms of final output, the choice of which database to use should be driven by other reason.
  One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see also Discussion). Another reason could be the frequency of updates and the potential for future growth, which prompted us to set \ac{ncbi} as the \ac{micone} standard for taxonomy assignment.
 In addition to being regularly maintained and updated the \ac{ncbi} database already has the advantage that its accuracy of assignments is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.

@@

  % TODO: Talk about the difference between correlations and associations
  The six different network inference methods used in this study are \ac{magma}~\cite{Cougoul2019}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
  Pairwise correlation methods involve calculating the correlation coefficient between every pair of \ac{otu}/\ac{esv}s leading to the detection of spurious indirect connections.
  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected \ac{otu}s~\cite{Kurtz2015,Menon2018}.

  For the analysis presented in this section, we used the taxonomy composition table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
  Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
  We can clearly see that the different networks differ vastly in their edge-density and connectivity and even some of the edges in common to these networks have their signs inverted. Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on [REF]).
  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they aren't intended for compositional datasets).
  The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
  The edge intersections in contrast show that only $19$ edges (out of $98$ edges in the smallest network - \ac{magma}) are in common between all the methods and each network has a large number of unique edges.
  These results indicate that there is a substantial rewiring of connections in the inferred networks.

@@

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{figure5.pdf}
  \end{figure}
  \begin{figure}[t!]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links are negative associations.
      A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
      \textbf{(B)} The node overlap Upset plot~\cite{Lex} indicates that all the networks have a large number of common nodes involved in connections.
      Whereas, \textbf{(C)} The edge overlap Upset plot shows that a very small fraction of these connections are actually shared.
    }
    \label{fig:figure5}
  \end{figure}
